The Galactic Midplane is Not a Plane: Implications for dynamical analysis with {\em Gaia} data and beyond

### 2019-04-22 paper/ms.tex
Add sentence about error distribution

\title{The Galactic Midplane is Not a Plane: Implications for {\em Gaia}}

@@

\section{Introduction} \label{sec:intro}
Our understanding of the Milky Way is currently undergoing a
revolution 
%mm in the era of {\em Gaia} DR2. 
     as a result of {\em Gaia} Data Release 2 (DR2).  Recent major discoveries include the remnants of a
major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the 2019arXiv190405721A
solar neighborhood \citep{2018Natur.561..360A} possibly indicating either
local substructure infall \citep{2018MNRAS.481.1501B, 2018arXiv180800451L} or
bar buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation
by a dark matter subtructure in the tidal stream GD1
\citep{2018ApJ...863L..20P, 2018arXiv181103631B}. These discoveries all
indicate that the Milky Way's stellar distribution, which demonstrably departs
significantly from axisymmetry, is undergoing phase mixing and dynamical
interactions across a range of spatial and temporal scales.

@@

In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
$J_z$ for the thick-disk orbit ({\em upper panel}) and thin-disk orbit ({\em
lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. the {\em right} panel
of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted
as a vertical dashed line. Here we see that the 
%mm spread 
    errors 
in $J_z$ induced by a
systematic offset in $z$ are non-Gaussian and bimodal. Furthermore, neither of
the modes lie %mm at the true value of $J_z$.
   at null. 
In the case of the thin-disk orbit
({\em lower panel}), we see that, in addition to the prior complications, the
error distribution is not even centered on 
%mm the true value. 
    null. 
This comes about
when the midplane error is $\gtrsim z_{\text{max}}$, and can be understood by
realizing that in such a case the midplane offset always acts to increase the
value of $J_z$, resulting in a biased estimate of $J_z$.

>>>>

\title{The Galactic Midplane is Not a Plane: Implications for dynamical analysis with {\em Gaia} data and beyond}

@@

\section{Introduction} \label{sec:intro}
Our understanding of the Milky Way is currently undergoing a
revolution 
%mm in the era of {\em Gaia} DR2. 
     as a result of {\em Gaia} Data Release 2 (DR2).  Recent major discoveries include the remnants of a
major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the
solar neighborhood \citep{2018Natur.561..360A} possibly indicating either
local substructure infall \citep{2018MNRAS.481.1501B, 2018arXiv180800451L} or
bar buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation
by a dark matter subtructure in the tidal stream GD1
\citep{2018ApJ...863L..20P, 2018arXiv181103631B}. These discoveries all
indicate that the Milky Way's stellar distribution, which demonstrably departs
significantly from axisymmetry, is undergoing phase mixing and dynamical
interactions across a range of spatial and temporal scales.

@@

In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
$J_z$ for the thick-disk orbit ({\em upper panel}) and thin-disk orbit ({\em
lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. the {\em right} panel
of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted
as a vertical dashed line. Here we see that the 
%mm spread 
    errors 
in $J_z$ induced by a
systematic offset in $z$ are non-Gaussian and bimodal. Furthermore, neither of
the modes lie %mm at the true value of $J_z$.
   at zero. In other words, neither of the modes of the distribution of observed $J_z$ values are centered on the true $J_z$ (i.e. the $J_z$ computed with no offset in $z$).
In the case of the thin-disk orbit
({\em lower panel}), we see that, in addition to the prior complications, the
error distribution is not even centered on 
%mm the true value. 
    null. 
This comes about
when the midplane error is $\gtrsim z_{\text{max}}$, and can be understood by
realizing that in such a case the midplane offset always acts to increase the
value of $J_z$, resulting in a biased estimate of $J_z$.
